Structured Versus Decorated Cospans

### 2022-03-10 ResponseToReviews.tex

{\bf We have moved the natural isomorphism triangle as equation (1) in the statement of Theorem 2.1, and we have referred to it at page 10.} {\chris 
$\checkmark$} {\kenny $\checkmark$}

\item p. 9, ingredients for constructing a double functor of decorated cospan double categories.

Does the natural transformation $\theta \colon EF \Rightarrow F’ H$ have to be a natural isomorphism? Quite possibly I missed something, but I don’t see where in 
the following proofs that assumption is used.

{\bf We thank the referee for catching this important mistake!   It does not need to be an isomorphism.
In fact, in Section 6 we construct a double functor
\[      \graysquare \colon F \lCsp \to D \lCsp \]
sending any open Petri net with rates to its corresponding open dynamical system, and we do this
using a monoidal natural transformation $\theta$ that is \emph{not} an isomorphism.  We had mistakenly written that it was; we have fixed this now.}   {\john: $\checkmark$}

@@

{\chris I see. Fixed, and see above.}
\fi

\item p. 27, end of Section 6.

It seems that the left adjoint needed for structured cospans fails to exist because the chosen category of dynamical systems does not have enough 
morphisms. Relatedly, this is the only example where the pseudofunctor D: (FinSet,+) $\to$ (Cat, $\times$) comes from a simple functor D: (FinSet,+) 
$\to$ (Set, $\times$) by viewing sets as discrete categories. In the other examples, this would count as using decorated cospans in the old, flawed 
way. 
All of this suggests the question: is there a way of constructing a category of dynamical systems that has more morphisms, so that a connection with 
structured cospans can be established?

{\bf That's a great question.   We do not know how to construct such a category, but the referee
is correct in suggesting that doing it might give a structured cospan category of open dynamical
systems.

As the referee noticed, the ``old, flawed'' construction of 
decorated cospans from a lax monoidal functor $D \colon (\mathsf{FinSet},+) \to 
\mathsf{Set}, \times) $ agrees with the new one using a lax monoidal pseudofunctor  $D \colon (\mathsf{FinSet},+) \to \mathsf{Cat}, \times) $ when the categories $D(S)$ are discrete,
and thus essentially sets.   Thus, our use of this construction amounts to working with a
set $D(S)$, rather than a full-fledged category, of dynamical systems on the finite set $S$.

In the example near the end of Section 6, the category $\int \! D$ can be seen as a category
of dynamical systems.   It is not discrete, but it does not have an initial object, as we'd
need to get a structured cospan
category that matches our decorated cospan category.

So, giving the category of dynamical systems more morphisms sounds like a good idea. We will continue thinking about it.}  { \john \checkmark}

@@

{\chris This looks similar to comment 30.}

{\john Yes, but it's worth a separate answer.   I wrote an answer above.  See if you like it.  }

@@

{\chris Right! We can perhaps add a sentence to this end, saying that although we only use E as an identity, someone could need to change decorations 
for some application?}

{\kenny Sure, maybe sometime before the official statement of the Theorem when we're describing how the double functor $\mathbb{H}$ is defined?}

{\chris Sounds good to me!}

\item Theorem 2.5: $F$ should be defined on $\mathsf{A}$, not $\mathsf{A}'$.

{\bf Done.} {\chris $\checkmark$} {\kenny $\checkmark$} {\john $\checkmark$}

@@

\item I think it would be worth saying in words a bit more prominently what the hypothesis of theorem 4.1 roughly means, namely that the fibers of 
the total 
category have finite colimits, reindexing preserves them, and the monoidal product on the fibers is cocartesian.

{\bf We have added a paragraph right below the statement of Theorem 4.1, clarifying its assumptions as suggested, which also incorporates an 
explanatory paragraph previously located under Corollary 4.3.} {\chris $\checkmark$} {\kenny $\checkmark$} {\john $\checkmark$}

### 2022-03-10 cospans_new.tex
more changes

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F, \maps (\A,+) \to (\Cat,\times)$ and $F' \maps (\A',+) \to 
(\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and 
a monoidal natural transformation $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\scriptstyle\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.  If $F, F'$ and $E$ are symmetric, then $\lH \maps F\lCsp \to F'\lCsp$ is a symmetric monoidal double functor. 
\end{thm}

>>>>

\begin{thm}
\label{thm:functoriality}
Given categories $\A$ and $\A'$ with finite colimits, lax monoidal pseudofunctors $F \maps (\A,+) \to (\Cat,\times)$ and $F' \maps (\A',+) \to 
(\Cat,\times)$, a finite colimit preserving functor $H \maps \A \to \A'$, a lax monoidal pseudofunctor $E \maps (\Cat,\times) \to (\Cat,\times)$ and 
a monoidal natural transformation $\theta$ as in the following diagram:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\Cat$};
\node (C) at (0,-1) {$\A'$};
\node (D) at (1,-1) {$\Cat$};
\node (E) at (0.5,-0.5) {$\scriptstyle\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(A) edge node[left]{$H$} (C)
(B) edge node[right]{$E$} (D)
(C) edge node[below]{$F'$} (D);
\end{tikzpicture}
\] 
we obtain a double functor $\lH \maps F\lCsp \to F'\lCsp$ defined as above.  If $F, F'$ and $E$ are symmetric, then $\lH \maps F\lCsp \to F'\lCsp$ is a symmetric monoidal double functor. 
\end{thm}
